Clusterization Properties of D-optimal Designs in Infinite Dimensional Bayesian Linear Inverse Problems

### 2023-04-17 latex/vanishing.tex
get ready to create different branches for different journals

\begin{proof}
  The proof amounts to using Woodbury's matrix identity twice and a
  regularization trick. The standard proof for Woodbury's matrix
  identity works in infinite dimensions, as long as all terms are well
  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
  it to be. Recall Woodbury's matrix identity and denote:
  \begin{align*}
    (A + UCV)^{-1} &= A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1} \\
    A :&= \prcov^{-1} \\
    U :&= \fwd^* \\
    V :&= \fwd \\
    C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
  \end{align*}

  Then, by Woodbury's matrix identity:
  \begin{equation}\label{eq:first}
  \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
      %
      %
      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
  %%   \end{split}
  %% \end{align*}
  \end{equation}
  Now, denote:
  \begin{align*}
    X :&= \fwd\prcov \fwd^* \\
    Y_{\eps} :&= \obs^*\obs + \eps I,
  \end{align*}
  and observe that $Y_{\eps}$ is invertible. Subsitute $X, Y_{\eps}$
  into the RHS of \eqref{eq:first}:
   \begin{equation}\label{eq:second}
     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
   \end{equation}

   Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
   two positive definite operators. Now, let
   \begin{align*}
     A :&= X^{-1}, \\
     C :&= \sigma^2Y_{\eps}^{-1}, \\
     U :&= I, \\
     V :&= I.
   \end{align*}
   Apply Woodbury's matrix identity to the RHS of \eqref{eq:second}:
  \begin{align*}
    \begin{split}
      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
      %
      %
      %
      &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
      %
      %
      %
      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
    \end{split}
  \end{align*}

@@

\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
  $\func$ acting as a linear functional. If
  \begin{equation*}
   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
  \end{equation*}
  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
  diagonalizable.
\end{lemma}

\begin{proof}
  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the indices of
  the eigenvectors corresponding to $\xi_i$
  \begin{equation*}
    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
  \end{equation*}
  Define further
  \begin{equation*}
    A_i := \sum_{k \in S_i} \func_k \func_k^*,
  \end{equation*}
  which is self-adjoint. Two observations are in order. First,
  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
  eigenvalue are orthogonal. For $k \in S_i$
  \begin{equation}\label{eq:on vi}
    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
  \end{equation}
  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
    vi}. A second application of \eqref{eq:on vi} shows that $A_i =
  \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and $C$ are
  simultaneously diagonalizable on $V_i$. This holds for every $1 \leq
  i \leq \ell$ and we conclude that $C$ and $A$ are simultaneously
  diagonalizable.
\end{proof}

\begin{proposition}[Eigenvectors of a D-optimal design]\label{prop:same ev}
  Let $\obs$ satisfy the nonlinear eigenvalue problem
  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ and $\fwd \prcov \fwd^*$
  are simultaneously diagonalizable.
\end{proposition}
\begin{proof}
  \begin{align}\label{eq:mod conditions}
    \begin{split}
      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
      %
      %
      %
      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
      %
      %
      %
      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
    \end{split}
  \end{align}

@@

\begin{proposition}\label{prop:true target}
  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
  k$ and $\eta_i = 0$ for $i > k$. Then:
  \begin{enumerate}
    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
      eigenvalues.
    \item
      \begin{equation*}
        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
      \end{equation*}
    \item Let $\obs$ a D-optimal design. Then $\eta_i > 0$ for
      eigenvectors corresponding to the $k$ largest $\lambda_i$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  Part (1) is trivial. To see part (2) holds: 
  \begin{align}
    \begin{split}
      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
      %
      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
      Theorem)}\\
      %
      %
      %
      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
      %
      %
      %
      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
      %
      %
      %
      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
      %
      %
      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
    \end{split}
  \end{align}
  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
\end{proof}

\begin{proposition}\label{prop:kkt}
  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\tar(\eta) =
  \frac{1}{2}\sum_{i=1}^m \log (1+\sigma^{-2}\lambda_i \eta_i)$, with
  $\lambda_i > 0$ and $\sigma^{2} > 0$. Then the maximum of $\tar$
  subject to $\eta_i \geq 0$ and $\sum\eta_i = m$ is obtained at
  \begin{equation}
  \eta_i = \begin{cases}
    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j\in A} \lambda_j^{-1} & i \in A \\
    0 & i \in A^c
  \end{cases}
  \end{equation}
  where $A:= \{1\leq i \leq m: \eta_i > 0\}$ and $A^c = \{1,\dots, m\}
    \backslash A$, and $|A|$ is the cardinality of $A$.
\end{proposition}

@@

We require Lemma \ref{lemma:free} for the proof of Theorem
\ref{thm:char}:

@@

Theorem \ref{thm:char} gives us a way to understand D-optimal designs
when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
corresponds to a graduated lab cylinder (Figure
\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
liquid of $\lambda_i^{-1}$ volume units. We are allowed $m$
observations, so we have blue liquid of volume $\sigma^{-2}m$ units at
our disposal. When we seek a D-optimal design, we distribute the blue
liquid by repeatedly adding a drop to whatever cylinder currently has
the lowest level of liquid in it, as long as its index $i \leq m$. The
result of such a procedure is illustrated in Figure \ref{fig:optimal
  vs not}.

@@

\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
The necessary conditions for D-optimality are (1), (2) and (4) of
Theorem \ref{thm:char}. As long as these conditions are satisfied,
$\obs$ is an optimizer of $\tar$. These conditions, however, still
leave some freedom in choosing $\obs$. This freedom may give rise to
clusterization, as illustrated in Figure \ref{fig:clusterization} and
demonstrated below.

Consider an inverse problem for which $\lambda_1^{-1} = 0.1,
\lambda_2^{-1} = 0.2,\sigma^2=1$ and $\{\ev_i\}_{i=1}^\infty$ are
eigenvectors of $\fwd\prcov\fwd^*$. Two D-optimal designs $\obs_1,
\obs_2$ presented in \eqref{eq:two designs}. It is easy to verify that
$\obs_1^*\obs_1 = \obs_2^*\obs_2$. The former exhibits clusterization
($\meas_3 = \meas_5 = \ev_1$), while the latter does not. Since both
of these designs are D-optimal, we conclude that clusterization poses
no obstruction to D-optimality.

@@

A computer implementation of Lemma \ref{lemma:free} (code in
Supplementary material) generates $\obs_1$ as a solution to the
D-optimal design problem mentioned in the previous
paragraph. Modifying the numerical values of the prior does not break
the clusterization. Thus, the preference to clusterization, at least
in Lemma \ref{lemma:free} is robust. The vast majority of randomized
numerical simulations (see code in Supplementary material) also give
rise to D-optimal designs that exhibit clustrization. We conclude that
clusterization poses no obstruction to D-optimality in the model
presented in this study. Moreover, sensor clusterization is generic,
as far as the model presented in this study is concerned.

%% Of course, a real-life problem is more restricted than the model we
%% consider here. Such restriction is the likely cause of clusterization
%% obseved in, e.g., the toy model of section
%% \ref{subsec:example}. Exactly how this happens is not clear. At this
%% point it is the subject of speculation and should be further studied
%% in the future.

>>>>

\begin{proof}
  The proof amounts to using Woodbury's matrix identity twice and a
  regularization trick. The standard proof for Woodbury's matrix
  identity works for separable Hilbert spaces, as long as all terms
  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
  we force it to be. Recall Woodbury's matrix identity:
  \begin{equation}\label{eq:WMI}
    (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}. 
  \end{equation}

  Denote:
  \begin{align}\label{eq:notation}
    \begin{split}
      A :&= \prcov^{-1} \\
      U :&= \fwd^* \\
      V :&= \fwd \\
      C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
    \end{split}
  \end{align}

  Then, by \eqref{eq:WMI}, with the notation \eqref{eq:notation}:
  \begin{equation}\label{eq:first}
  \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
      %
      %
      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
  %%   \end{split}
  %% \end{align*}
  \end{equation}
  Now, denote:
  \begin{align*}
    X :&= \fwd\prcov \fwd^* \\
    Y_{\eps} :&= \obs^*\obs + \eps I,
  \end{align*}
  and observe that $Y_{\eps}$ is invertible. Subsitute $X, Y_{\eps}$
  into the RHS of \eqref{eq:first}:
   \begin{equation}\label{eq:second}
     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
   \end{equation}

   Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
   two positive definite operators. Now, let
   \begin{align}\label{eq:notation2}
     \begin{split}
       A :&= X^{-1}, \\
       C :&= \sigma^2Y_{\eps}^{-1}, \\
       U :&= I, \\
       V :&= I.
     \end{split}
   \end{align}
   Apply \eqref{eq:WMI} with the notation \eqref{eq:notation2} to the
   RHS of \eqref{eq:second}:
  \begin{align*}
    \begin{split}
      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
      %
      %
      %
      &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
      %
      %
      %
      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
    \end{split}
  \end{align*}

@@

\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
  $\func$ acting as a linear functional. If
  \begin{equation*}
   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
  \end{equation*}
  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
  diagonalizable.
\end{lemma}
\begin{proof}
  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the indices of
  the eigenvectors corresponding to $\xi_i$
  \begin{equation*}
    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
  \end{equation*}
  Define further
  \begin{equation*}
    A_i := \sum_{k \in S_i} \func_k \func_k^*,
  \end{equation*}
  which is self-adjoint. Two observations are in order. First,
  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
  eigenvalue are orthogonal. For $k \in S_i$
  \begin{equation}\label{eq:on vi}
    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
  \end{equation}
  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
    vi}. A second application of \eqref{eq:on vi} shows that $A_i =
  \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and $C$ are
  simultaneously diagonalizable on $V_i$. This holds for every $1 \leq
  i \leq \ell$ and we conclude that $C$ and $A$ are simultaneously
  diagonalizable.
\end{proof}

\begin{proposition}\label{prop:same ev}
  Let $\obs$ satisfy the nonlinear eigenvalue problem
  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ and $\fwd \prcov \fwd^*$
  are simultaneously diagonalizable.
\end{proposition}
\begin{proof}
  \begin{align}\label{eq:mod conditions}
    \begin{split}
      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
      %
      %
      %
      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
      %
      %
      %
      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
    \end{split}
  \end{align}

@@

\begin{proposition}\label{prop:true target}
  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
  k$ and $\eta_i = 0$ for $i > k$. Then:
  \begin{enumerate}
    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
      eigenvalues.
    \item
      \begin{equation*}
        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
      \end{equation*}
    \item Furthermore, if $\obs$ isD-optimal, $\eta_i > 0$ for
      eigenvectors corresponding to the $k$ largest $\lambda_i$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  Part (1) is trivial. To see part (2) holds: 
  \begin{align}
    \begin{split}
      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
      %
      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
      Theorem)}\\
      %
      %
      %
      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
      %
      %
      %
      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
      %
      %
      %
      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
      %
      %
      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
    \end{split}
  \end{align}
  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
\end{proof}

\begin{proposition}\label{prop:kkt}
  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\tar(\eta) =
  \frac{1}{2}\sum_{i=1}^m \log (1+\sigma^{-2}\lambda_i \eta_i)$, with
  $\lambda_i > 0$ and $\sigma^{2} > 0$. Then the maximum of $\tar$
  subject to $\eta_i \geq 0$ and $\sum\eta_i = m$ is obtained at
  \begin{equation}
  \eta_i = \begin{cases}
    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j\in A} \lambda_j^{-1} & i \in A \\
    0 & i \in A^c
  \end{cases}
  \end{equation}
  where $A:= \{1\leq i \leq m: \eta_i > 0\}$ and $A^c = \{1,\dots, m\}
  \backslash A$, and $k = |A|$, the cardinality of $A$.
\end{proposition}

@@

The final ingredient we require for the proof of Theorem
\ref{thm:char} the following lemma:

@@

Theorem \ref{thm:char} facilitates understanding of D-optimal designs
when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
corresponds to a graduated lab cylinder (Figure
\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
liquid of $\lambda_i^{-1}$ volume units. We are allowed $m$
observations, so we have blue liquid of volume $\sigma^{-2}m$ units at
our disposal. When we seek a D-optimal design, we distribute the blue
liquid by repeatedly adding a drop to whatever cylinder currently has
the lowest level of liquid in it, as long as its index $i \leq m$. The
result of such a procedure is illustrated in Figure \ref{fig:optimal
  vs not}.

@@

\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
The necessary conditions for D-optimality are (1), (2) and (4) of
Theorem \ref{thm:char}. These conditions, however, still leave some
freedom in choosing $\obs$. This freedom may give rise to
clusterization, as illustrated in Figure \ref{fig:clusterization} and
demonstrated below.

Consider an inverse problem for which $\lambda_1^{-1} = 0.1,
\lambda_2^{-1} = 0.2,\sigma^2=1$ and $\{\ev_i\}_{i=1}^\infty$ are
eigenvectors of $\fwd\prcov\fwd^*$. Two D-optimal designs $\obs_1,
\obs_2$ are presented in \eqref{eq:two designs}, and it is easy to
verify that $\obs_1^*\obs_1 = \obs_2^*\obs_2$. However, $\obs_1$
exhibits clusterization ($\meas_3 = \meas_5 = \ev_1$), while $\obs_2$
does not. Both of $\obs_1$ and $\obs_2$ are D-optimal, and we conclude
that clusterization poses no obstruction to D-optimality.

@@

A computer implementation of Lemma \ref{thm:char} (code in
Supplementary material) generates $\obs_1$ as a solution to the
D-optimal design problem mentioned in the previous
paragraph. Randomized numerical simulations (see code in Supplementary
material) give rise to D-optimal designs that exhibit clustrization
more than 95\% of the time. We conclude that clusterization poses no
obstruction to D-optimality for the model presented in this study and
that sensor clusterization is generic.
